### 2021-06-28 notes/notes.tex
Added bullet points for results

\end{tabularx}
\end{chapter}

@@

\cite{Smith_Interacting_2006} speak of error-clamp weirdness and presents a
model to explain it. Their data is limited, however, and might not be specially
useful for me.

\cite{Ethier_Spontaneous_2008} have error-clamp weirdness and their
within-block pauses might be useful for my interpretations.
\end{chapter}

@@

\cite{Vaswani_Decay_2013} explores in detail the motor decay observed during EC
trials after adaptation trials. They conclude, among other things, that the
motor adaptation decays across as many as hundreds of trials, but never goes
back to zero. Additionally, there is a lag before decay starts (all their
experiments end in A-E, so no baseline before), and it depends on the
participant, as well as how noticeable the transition to error-clamp trials
is. They have some clever experiments.

@@

@@

\begin{center}
\begin{tabularx}{0.8\textwidth}{ c | c | X}
  Variable & Value & Description \\ \hline
  prediction\_noise & 3 & The bigger this is, the slower correct inference is. Only with no cues, though. \\  
  cue\_noise & 0.5 & Creates uninformative cues.
  force\_sds & 0.0002 & Changing this makes the model learn more quickly or slowly. With this value, the agent barely finishes adaptation by the time the adaptation block is over.
  context\_noise & 0 & Higher numbers make context inference more uncertain after evidence accumulation.  
\end{tabularx}
\end{center}

@@

>>>>

\section{Results}

\begin{enumerate}
\item Quick de-adaptation/savings: gradual evidence accumulation during context inference will make the switching between internal models not be immediate.
\item Learning: The speed in which the internal models are adjusted given errors depends on how certain the inference is. More uncertain context inference will lead to slower learning.
\item Action selection: Action selection is sampled from the existing (relevant) internal models, weighted by their posterior probability during this trial.
\item Cue reliability adjustments: As more bad-cue trials are observed, the agent will lower the estimated reliability of the cue.
\item Error-clamp tappering off: Hundreds of error-clamp trials make adaptation seem to slowly disappear, but it never fully goes away.
\item Error-clamp lag: For a while, the tappering off does not begin, depending on how clear it is that error-clamp trials have started.
\end{enumerate}

@@

\cite{Smith_Interacting_2006} present a typical mechanical arm task with a
O-A-(-A)-E structure, with random E trials thrown in throughout all. Half of
the trials were with outward movements and half with the inward movement, with
the latter almost being E trials and not used for analysis. The experiment has
no cues, but it was a design in which blocks of 60 trials were used, presumably
with pauses in the middle, though they do not explicitly say this. The authors
find that during the E block at the end, participants' responses show
spontaneous recovery, where they seem to be moving towards A for a few trials,
but never quite reaching it, and then taper off to
zero. \cite{Smith_Interacting_2006} only show data aggregated over all
participants, so it's difficult to extract too much information, but their
figure 3D suggests that some participants assume -A during E, while most assume
A.

\cite{Ethier_Spontaneous_2008} present a saccade study (movements of 15° with
adaptation of +-5°) in which participants go through E-A-(-A)-E blocks, each
divided into sets of 60 trials (some blocks have different sets). Between each
set (and block), a pause of 30s happened. There are no cues, although the size
of the adaptations is big. It is important to note that the way the error clamp
works in this paradigm probably gives no indication to the participant that an
error has occured; contrast this with mechanical-arm paradigms, or
visuorotation reaching paradigms, in which proprioceptive information gives
feedback to the participant regardless of rewards.

Figure 2 in \cite{Ethier_Spontaneous_2008} contains all the results relevant to
this section. First of all, it is clear from the first E block that spontaneous
recovery and slow tapering off do not exist if no adaptation has been learned;
instead, behavior during these trials is the same as during baseline. Second,
spontaneous recovery happens always in the direction of the first adaptation A,
and never in the direction of -A, which is the block right before the last E
block. The context-inference-based model explains these findings by noting that
the -A block is very short, ending when participants reach 0 adaptation (during
``de-adaptation'') or a little bit beyond zero (towards -A). This is important
because, since adaptation was not finished, during the first trial of the E
block the participant expected an error, but no error was observed. This leads
the participant to think that a change in the context has occured, and context
inference kicks in, taking the lack of an error as evidence in favor of either
baseline or the previous adaptation A. Given that the action by the participant
on this trial is more consistent with baseline than with A, why would A play
any role in future responses? The answer lies in the structure of the
experiment, as no baseline trials were ever performed in front of the screen,
which makes baseline have very low priors.

@@

\cite{Vaswani_Decay_2013} explores in detail the motor decay observed during EC
trials after adaptation trials in a viscous field paradigm. They conclude,
among other things, that the motor adaptation decays across as many as hundreds
of trials, but never goes back to zero. Additionally, there is a lag before
decay starts (all their experiments end in A-E, so no baseline before), and it
depends on the participant, as well as how noticeable the transition to
error-clamp trials is. They have some clever experiments.

More precisely, \cite{Vaswani_Decay_2013} have an experiment 1 in which
participants learn an adaptation and then go through hundreds of trials of
error clamp. Participants are divided into four groups: groups 1.1 and 1.4
learn adaptation A and -A, respectively, and then go to EC trials. Group 1.2
performs the task in baseline and then in A. Group 1.3 performs the task in
-A/2, then A and then EC. In their results, all groups show ``adaptation'' that
starts high (consistent with A) and, after some lag, goes down to an adaptation
of almost zero, but decidedly positive. Group 1.3, in particular, decays faster
than the other groups but does not seem to get closer to zero even at the end
of the experiment. Group 1.4 is just like 1.1 and 1.2, but with its sign
flipped.

Our simulations can easily show these results to be a direct consequence of
context inference, which starts by assuming that no change has happened after
EC starts (thus staying with A), but eventually the evidence points towards a
change. However, because no model explains EC fully, actions are taken by doing
a weighted average of all available models, where the weights are given by the
posterior over contexts. This explains both the lag, the fact that EC
adaptation never reaches zero and that group 1.3 adapted faster, as the -A/2
model pulls the average down faster than baseline. The fact that no difference
is observed between 1.1 and 1.2 is due to the fact that the baseline model
exists, regardless of whether participants trained on it or not. Just like the
results from \cite{Davidson_Scaling_2004}, these results point toward motor
output being a weighted average of all the available models, with weights being
the posterior over contexts.

The conclusions of experiments 2 and 3 by \cite{Vaswani_Decay_2013} reinforce
the context-inference-based account: the more difficult it is to notice that EC
has started, the less decay participants show, to the extreme of experiment 3
in which most participants showed no decay at all.

\cite{Pekny_Protection_2011} present a series of experiments in which they set
out to determine whether A-B-E paradigms make the brain forget about A when
learning B, or whether the brain protects A. The authors mention that there is
a lot of literature that has used savings as a way to measure the protection of
A and found that A is not being protected; however, the authors argue that
classical conditioning experiments instead use spontaneous recovery during E
trials to measure protection of memories during extinction, and the literature
in that field has demonstrated that memories are indeed
protected. \cite{Pekny_Protection_2011} present a series of experiments with
mechanical arms to show the protection of memories. Their results, specifically
those in their Figures 3 and 5, strongly support not only the
context-inference-based motor learner, but also the idea that action selection
happens through a weighted average, much like the results from
\cite{Herzfeld_memory_2014}.

In figure 3A, one group (red) went through O-A-b (where lowercase ``b'' means
very few trials, used as extinction), and another group (black) went through
O-B-A-b. When comparing the force compensation between those two groups, one
can see that the introduction of B in the black group has the effect of pulling
their response during E closer to baseline, as compared with the red
group. This is because B has, for this group, been introduced as a potential
context of importance equal that of A (although a longer time ago, as B is
presented before A). This increased the prior probability of context B during
the E trials in the minds of participants, which draws the weighted average
closer to baseline, as $B = -A$. The same interpretation can be given to all of
Figure 3 and Figure 5 as well.

Additionally, \cite{Pekny_Protection_2011} present an experiment in which
changes in context are presented gradually (see their figure 1 for
details). They argue that because the changes are presented gradually,
participants do not get a strong prediction error that can guide context
inference towards the correct context. However, there is a caveat they do not
discuss: in an experiment where context switches slowly, small changes become
big evidence. For this reason, their Figure 5B has the same effect as that seen
in Figure 3.

@@

\end{chapter}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{center}
\begin{tabularx}{0.8\textwidth}{ c | c | X}
  Variable & Value & Description \\ \hline
  prediction\_noise & 3 & The bigger this is, the slower correct inference is. Only with no cues, though. \\  
  cue\_noise & 0.5 & Creates uninformative cues. \\
  force\_sds & 0.0002 & Changing this makes the model learn more quickly or slowly. With this value, the agent barely finishes adaptation by the time the adaptation block is over. \\
  context\_noise & 0 & Higher numbers make context inference more uncertain after evidence accumulation.  
\end{tabularx}
\end{center}

@@

\subsection{\cite{Davidson_Scaling_2004} results}
Running simulations I realized that the model, as is, will not be able to
reproduce the results from \cite{Davidson_Scaling_2004}. To do so, the model
would need to be modified to create new models online, but only one at a
time. A model would be created and kept at an adaptation of zero with very high
standard deviation, and this model would very naturally learn any adaptation
that is not explained by any other model.

The biggest problem is in reproducing experiment 2, group 1 of
\cite{Davidson_Scaling_2004}, in which the adaptations that will be presented
are A and 3A. In this case, at the beginning there should be an empty model
that learns A. Once this model is far away from baseline, a new empty model
should be created (and kept ``on call''), which eventually would pick up on
3A. If the two empty models are created at the same time, both will learn A,
and later only one of them (picked randomly) will learn 3A, while the other
stays at A; this is fine for simulations but difficult to explain from a
mechanistic point of view: why would the brain train two models to do the same
thing?

Possible solutions to this would be: (1) as said before, that only one empty
model exists at any time. (2) any model would have an empty counterpart that
can always learn faster. (3) Context inference would be able to pick up on low
model evidence (e.g. when no internal model explains the observations at all)
and trigger the creation of a new model, either always as a copy of baseline or
as a copy of the closest to the observations.

This is work for future studies.
